No time for dead time - Use the Fourier amplitude differences to normalize dead time-affected periodograms

### 2017-11-01 paper/deadtime.tex
Some modifications to start addressing referee's comments and allow further clarifications

\section{Data simulation and analysis} \label{sec:data}
\subsection{Simulated datasets}
All simulated data sets in this paper have been produced with a combination of the two Python libraries \texttt{stingray} \citep{huppenkothen2016} and \texttt{HENDRICS} \citep[formerly known as MaLTPyNT][]{2015ascl.soft02021B}, both based on Astropy \citep{astropy2013}.
The \texttt{stingray.simulate.Simulator} class was used to simulate light curves with a given noise profile, 
the \texttt{stingray.Eventlist.simulate\_times()} function was used to transform the light curves into event lists using rejection sampling, 
and finally the \texttt{hendrics.fake.filter\_for\_deadtime()} function was used to apply a non-paralyzable dead time to the simulated event lists. For more details on the simulated data sets, see also Section \ref{sec:correction} and the available Jupyter notebooks\footnote{\href{https://github.com/matteobachetti/deadtime-paper-II}{https://github.com/matteobachetti/deadtime-paper-II}} \citep[for a description of Jupyter notebooks, see][]{kluyver2016jupyter}.

\subsection{Cyg X-1 \nustar dataset}
We downloaded the observation directory of ObsID 30001011009 (UT 2014-10-04) from the HEASARC using the custom \texttt{heasarc\_pipelines} package [ASCL in prep.]. 
Starting from the standard cleaned science event files distributed by HEASARC, we applied a barycenter correction using the FTOOL \texttt{barycorr} shipped with HEASOFT 6.21 with the clock correction file n. 71 from the \nustar CALDB.
We selected photons in a region of 50\arcsec around the nominal position of Cyg X-1 using the FTOOL \texttt{fselect}.
We used the scripts contained in \texttt{HENDRICS} to load the event lists for the two detectors FPMA and FPMB, calibrate them to translate the PI channels into energy values, and selected photons from 3 to 79 keV.

@@

\section{dead time-affected white noise} \label{sec:wndeadtime}

@@

The correction is depicted in Figure~\ref{fig:pds}, right. 
Starting from a heavily distorted distribution of the powers, applying the FAD correction reinstates a remarkably correct distribution of powers, following the expected $\chi^2_2$ distribution \citep{Lewin+88} very closely.
Analogously, the corrected cospectrum will follow the expected Laplace distribution.
While the original dead time-affected cospectrum had a frequency-dependent modification to the rms level, the FAD-corrected cospectrum gets back to a frequency-independent shape, like in the dead time-free case.

@@

\bibliographystyle{aasjournal}
%\bibliography{deadtime,papers3}
\begin{thebibliography}{}
\expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi
\providecommand{\url}[1]{\href{#1}{#1}}

\bibitem[{{Astropy Collaboration} {et~al.}(2013){Astropy Collaboration},
  {Robitaille}, {Tollerud}, {Greenfield}, {Droettboom}, {Bray}, {Aldcroft},
  {Davis}, {Ginsburg}, {Price-Whelan}, {Kerzendorf}, {Conley}, {Crighton},
  {Barbary}, {Muna}, {Ferguson}, {Grollier}, {Parikh}, {Nair}, {Unther},
  {Deil}, {Woillez}, {Conseil}, {Kramer}, {Turner}, {Singer}, {Fox}, {Weaver},
  {Zabalza}, {Edwards}, {Azalee Bostroem}, {Burke}, {Casey}, {Crawford},
  {Dencheva}, {Ely}, {Jenness}, {Labrie}, {Lim}, {Pierfederici}, {Pontzen},
  {Ptak}, {Refsdal}, {Servillat}, \& {Streicher}}]{astropy2013}
{Astropy Collaboration}, {Robitaille}, T.~P., {Tollerud}, E.~J., {et~al.} 2013,
  \aap, 558, A33

\bibitem[{Bachetti(2015)}]{2015ascl.soft02021B}
Bachetti, M. 2015, Astrophysics Source Code Library, ascl:1502.021

\bibitem[{Bachetti {et~al.}(2015)Bachetti, Harrison, Cook, Tomsick, Schmid,
  Grefenstette, Barret, Boggs, Christensen, Craig, Fabian, F{\"u}rst, Gandhi,
  Hailey, Kara, Maccarone, Miller, Pottschmidt, Stern, Uttley, Walton, Wilms,
  \& Zhang}]{Bachetti+15}
Bachetti, M., Harrison, F.~A., Cook, R., {et~al.} 2015, ApJ, 800, 109

\bibitem[{Barret \& Vaughan(2012)}]{Barret+12}
Barret, D., \& Vaughan, S. 2012, ApJ, 746, 131

\bibitem[{Belloni \& Hasinger(1990)}]{BelloniHasinger90}
Belloni, T., \& Hasinger, G. 1990, A{\&}A, 230, 103

\bibitem[{Bowyer {et~al.}(1965)Bowyer, Byram, Chubb, \& Friedman}]{Bowyer+65}
Bowyer, S., Byram, E.~T., Chubb, T.~A., \& Friedman, H. 1965, Astronomical
  Observations from Space Vehicles, 23, 227

\bibitem[{Gilfanov {et~al.}(2000)Gilfanov, Churazov, \&
  Revnivtsev}]{Gilfanov+00}
Gilfanov, M., Churazov, E., \& Revnivtsev, M. 2000, MNRAS, 316, 923

\bibitem[{Grinberg {et~al.}(2013)Grinberg, Hell, Pottschmidt, B{\"o}ck, Nowak,
  Rodriguez, Bodaghee, Cadolle~Bel, Case, Hanke, K{\"u}hnel, Markoff, Pooley,
  Rothschild, Tomsick, Wilson-Hodge, \& Wilms}]{Grinberg+13}
Grinberg, V., Hell, N., Pottschmidt, K., {et~al.} 2013, 554, 88

\bibitem[{{Huppenkothen} {et~al.}(2016){Huppenkothen}, {Bachetti}, {Stevens},
  {Migliari}, \& {Balm}}]{huppenkothen2016}
{Huppenkothen}, D., {Bachetti}, M., {Stevens}, A.~L., {Migliari}, S., \&
  {Balm}, P. 2016, {Stingray: Spectral-timing software}, Astrophysics Source
  Code Library, , , ascl:1608.001

\bibitem[{Huppenkothen {et~al.}(2017)Huppenkothen, Younes, Ingram, Kouveliotou,
  G{\"o}{\u{g}}{\"u}{\c s}, Bachetti, Sanchez-Fernandez, Chenevez, Motta,
  van~der Klis, Granot, Gehrels, Kuulkers, Tomsick, \&
  Walton}]{Huppenkothen+17}
Huppenkothen, D., Younes, G., Ingram, A., {et~al.} 2017, ApJ, 834, 90

\bibitem[{Kluyver {et~al.}(2016)Kluyver, Ragan-Kelley, P{\'e}rez, Granger,
  Bussonnier, Frederic, Kelley, Hamrick, Grout, Corlay,
  {et~al.}}]{kluyver2016jupyter}
Kluyver, T., Ragan-Kelley, B., P{\'e}rez, F., {et~al.} 2016, in ELPUB, 87--90

\bibitem[{Leahy {et~al.}(1983)Leahy, Darbro, Elsner, Weisskopf, Kahn,
  Sutherland, \& Grindlay}]{Leahy+83}
Leahy, D.~A., Darbro, W., Elsner, R.~F., {et~al.} 1983, ApJ, 266, 160

\bibitem[{Lewin {et~al.}(1988)Lewin, van Paradijs, \& van~der Klis}]{Lewin+88}
Lewin, W. H.~G., van Paradijs, J., \& van~der Klis, M. 1988, SSRv, 46, 273

\bibitem[{Miyamoto {et~al.}(1991)Miyamoto, Kimura, Kitamoto, Dotani, \&
  Ebisawa}]{Miyamoto+91}
Miyamoto, S., Kimura, K., Kitamoto, S., Dotani, T., \& Ebisawa, K. 1991, ApJ,
  383, 784

\bibitem[{Vikhlinin {et~al.}(1994)Vikhlinin, Churazov, \&
  Gilfanov}]{Vikhlinin+94}
Vikhlinin, A., Churazov, E., \& Gilfanov, M. 1994, 287, 73

\bibitem[{Zhang {et~al.}(1995)Zhang, Jahoda, Swank, Morgan, \&
  Giles}]{Zhang+95}
Zhang, W., Jahoda, K., Swank, J.~H., Morgan, E.~H., \& Giles, A.~B. 1995, ApJ,
  449, 930

\end{thebibliography}
\end{document}

>>>>

\section{Data simulation and analysis} \label{sec:data}
All simulated and real data sets in this paper were produced and/or analyzed with a combination of the two Python libraries \texttt{stingray}%
\footnote{The library is under heavy development.
    For this work we used the version identified by the hash \texttt{XXXXXX}}
\citep{huppenkothen2016} and \texttt{HENDRICS} v.\texttt{3.0b1} \citep[formerly known as MaLTPyNT;][]{2015ascl.soft02021B}, both based on Astropy \citep{astropy2013}.

\subsection{Simulated datasets}
We used the same procedure and algorithms described by \citet{Bachetti+15}, Section 4, that we briefly summarize here together with references to the methods and classes of the libraries above where these steps are implemented.
The classes in \texttt{stingray.simulate} were used to simulate light curves with a given noise profile. 
These classes use the \citet{timmer1995} method to create a noise-affected light curve starting from a given power spectral shape described by a powerlaw index, an \texttt{astropy} model or a list of arrays containing the frequency and power of the model spectrum.
We then used an acceptance-rejection method to obtain simulated events from these light curves (\texttt{stingray.Eventlist.simulate\_times()}). 
Finally, the \texttt{hendrics.fake.filter\_for\_deadtime()} function was used to apply a non-paralyzable dead time of 2.5\,ms to the simulated event lists. For more details on the simulated data sets, see also Section \ref{sec:correction} and the available Jupyter notebooks%
\footnote{\href{https://github.com/matteobachetti/deadtime-paper-II}{https://github.com/matteobachetti/deadtime-paper-II}} \citep[for a description of Jupyter notebooks, see][]{kluyver2016jupyter}.

\subsection{Cyg X-1 \nustar dataset}
We downloaded the observation directory of \nustar ObsID 30001011009 (UT 2014-10-04) from the HEASARC using the custom \texttt{heasarc\_pipelines} package (Bachetti et al., ASCL identifier requested).
Starting from the standard cleaned science event files distributed by HEASARC, we applied a barycenter correction using the FTOOL \texttt{barycorr} shipped with HEASOFT 6.21 with the clock correction file n. 71 from the \nustar CALDB.
We selected photons in a region of 50\arcsec around the nominal position of Cyg X-1 using the FTOOL \texttt{fselect}.
We used the scripts contained in \texttt{HENDRICS} to load the event lists for the two detectors FPMA and FPMB, calibrate them to translate the PI channels into energy values, and produced light curves with photons from 3 to 79 keV%
\footnote{This procedure uses the scripts \texttt{HENreadfile}, \texttt{HENcalibrate} and \texttt{HENlcurve}, as described at\\ \href{http://hendrics.readthedocs.io/en/master/tutorials/quicklook.html}{http://hendrics.readthedocs.io/en/master/tutorials/quicklook.html}}.

@@

\section{Dead time-affected white noise} \label{sec:wndeadtime}

@@

The correction is depicted in Figure~\ref{fig:dist}, right. 
Starting from a heavily distorted distribution of the powers, applying the FAD correction reinstates a remarkably correct distribution of powers, following the expected $\chi^2_2$ distribution \citep{Lewin+88} very closely.
Analogously, the corrected cospectrum will follow the expected Laplace distribution.
While the original dead time-affected cospectrum had a frequency-dependent modification to the rms level, the FAD-corrected cospectrum gets back to a frequency-independent shape, like in the dead time-free case.

@@

\bibliographystyle{aasjournal}
\bibliography{deadtime,papers3}
%\begin{thebibliography}{}
%\expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi
%\providecommand{\url}[1]{\href{#1}{#1}}
%
%\bibitem[{{Astropy Collaboration} {et~al.}(2013){Astropy Collaboration},
%  {Robitaille}, {Tollerud}, {Greenfield}, {Droettboom}, {Bray}, {Aldcroft},
%  {Davis}, {Ginsburg}, {Price-Whelan}, {Kerzendorf}, {Conley}, {Crighton},
%  {Barbary}, {Muna}, {Ferguson}, {Grollier}, {Parikh}, {Nair}, {Unther},
%  {Deil}, {Woillez}, {Conseil}, {Kramer}, {Turner}, {Singer}, {Fox}, {Weaver},
%  {Zabalza}, {Edwards}, {Azalee Bostroem}, {Burke}, {Casey}, {Crawford},
%  {Dencheva}, {Ely}, {Jenness}, {Labrie}, {Lim}, {Pierfederici}, {Pontzen},
%  {Ptak}, {Refsdal}, {Servillat}, \& {Streicher}}]{astropy2013}
%{Astropy Collaboration}, {Robitaille}, T.~P., {Tollerud}, E.~J., {et~al.} 2013,
%  \aap, 558, A33
%
%\bibitem[{Bachetti(2015)}]{2015ascl.soft02021B}
%Bachetti, M. 2015, Astrophysics Source Code Library, ascl:1502.021
%
%\bibitem[{Bachetti {et~al.}(2015)Bachetti, Harrison, Cook, Tomsick, Schmid,
%  Grefenstette, Barret, Boggs, Christensen, Craig, Fabian, F{\"u}rst, Gandhi,
%  Hailey, Kara, Maccarone, Miller, Pottschmidt, Stern, Uttley, Walton, Wilms,
%  \& Zhang}]{Bachetti+15}
%Bachetti, M., Harrison, F.~A., Cook, R., {et~al.} 2015, ApJ, 800, 109
%
%\bibitem[{Barret \& Vaughan(2012)}]{Barret+12}
%Barret, D., \& Vaughan, S. 2012, ApJ, 746, 131
%
%\bibitem[{Belloni \& Hasinger(1990)}]{BelloniHasinger90}
%Belloni, T., \& Hasinger, G. 1990, A{\&}A, 230, 103
%
%\bibitem[{Bowyer {et~al.}(1965)Bowyer, Byram, Chubb, \& Friedman}]{Bowyer+65}
%Bowyer, S., Byram, E.~T., Chubb, T.~A., \& Friedman, H. 1965, Astronomical
%  Observations from Space Vehicles, 23, 227
%
%\bibitem[{Gilfanov {et~al.}(2000)Gilfanov, Churazov, \&
%  Revnivtsev}]{Gilfanov+00}
%Gilfanov, M., Churazov, E., \& Revnivtsev, M. 2000, MNRAS, 316, 923
%
%\bibitem[{Grinberg {et~al.}(2013)Grinberg, Hell, Pottschmidt, B{\"o}ck, Nowak,
%  Rodriguez, Bodaghee, Cadolle~Bel, Case, Hanke, K{\"u}hnel, Markoff, Pooley,
%  Rothschild, Tomsick, Wilson-Hodge, \& Wilms}]{Grinberg+13}
%Grinberg, V., Hell, N., Pottschmidt, K., {et~al.} 2013, 554, 88
%
%\bibitem[{{Huppenkothen} {et~al.}(2016){Huppenkothen}, {Bachetti}, {Stevens},
%  {Migliari}, \& {Balm}}]{huppenkothen2016}
%{Huppenkothen}, D., {Bachetti}, M., {Stevens}, A.~L., {Migliari}, S., \&
%  {Balm}, P. 2016, {Stingray: Spectral-timing software}, Astrophysics Source
%  Code Library, , , ascl:1608.001
%
%\bibitem[{Huppenkothen {et~al.}(2017)Huppenkothen, Younes, Ingram, Kouveliotou,
%  G{\"o}{\u{g}}{\"u}{\c s}, Bachetti, Sanchez-Fernandez, Chenevez, Motta,
%  van~der Klis, Granot, Gehrels, Kuulkers, Tomsick, \&
%  Walton}]{Huppenkothen+17}
%Huppenkothen, D., Younes, G., Ingram, A., {et~al.} 2017, ApJ, 834, 90
%
%\bibitem[{Kluyver {et~al.}(2016)Kluyver, Ragan-Kelley, P{\'e}rez, Granger,
%  Bussonnier, Frederic, Kelley, Hamrick, Grout, Corlay,
%  {et~al.}}]{kluyver2016jupyter}
%Kluyver, T., Ragan-Kelley, B., P{\'e}rez, F., {et~al.} 2016, in ELPUB, 87--90
%
%\bibitem[{Leahy {et~al.}(1983)Leahy, Darbro, Elsner, Weisskopf, Kahn,
%  Sutherland, \& Grindlay}]{Leahy+83}
%Leahy, D.~A., Darbro, W., Elsner, R.~F., {et~al.} 1983, ApJ, 266, 160
%
%\bibitem[{Lewin {et~al.}(1988)Lewin, van Paradijs, \& van~der Klis}]{Lewin+88}
%Lewin, W. H.~G., van Paradijs, J., \& van~der Klis, M. 1988, SSRv, 46, 273
%
%\bibitem[{Miyamoto {et~al.}(1991)Miyamoto, Kimura, Kitamoto, Dotani, \&
%  Ebisawa}]{Miyamoto+91}
%Miyamoto, S., Kimura, K., Kitamoto, S., Dotani, T., \& Ebisawa, K. 1991, ApJ,
%  383, 784
%
%\bibitem[{Vikhlinin {et~al.}(1994)Vikhlinin, Churazov, \&
%  Gilfanov}]{Vikhlinin+94}
%Vikhlinin, A., Churazov, E., \& Gilfanov, M. 1994, 287, 73
%
%\bibitem[{Zhang {et~al.}(1995)Zhang, Jahoda, Swank, Morgan, \&
%  Giles}]{Zhang+95}
%Zhang, W., Jahoda, K., Swank, J.~H., Morgan, E.~H., \& Giles, A.~B. 1995, ApJ,
%  449, 930
%
%\end{thebibliography}
\end{document}
